Algebraic Groups\vspace{-1ex}

### 2020-03-07 Algebraic Geometry/Algebraic-Geometry.tex
Update because all the classes are ending

>>>>

\section{March 2nd, 2020}
Sorry, I missed Friday. Today we are continuing to talk about 
\subsection{Closed Subschemes}
We saw last time that there is a bijection between closed subschemes $\iota:Y\hookrightarrow X$ (modulo some relation?) and the collection of quasicoherent sheaves of ideals $\calI\subseteq\O_X$.
Last time Max showed that it is surjective.

To show injectivity, we go through a few steps: we begin by showing that the topological spaces underlying 
two closed immersions $Y$ and $Y'$ with an iso $g:\iota_\ast\O_Y\to \iota'_\ast\O_{Y'}$. Then we continued by 
constructing stuff.

\section{March 4th, 2020}
Today we are talking about sheaves on proj. The

### 2020-03-07 Algebraic Groups/Algebraic-Groups.tex
Update because all the classes are ending

% Document-Specific Macros
\DeclareMathOperator{\SL}{SL}

@@

\begin{prop}
	\begin{enumerate}
		\item The functor sending 
		\[M\mapsto H^0(X/G,\calL(M))=\Gamma(\calL(M))\]
		is left exact (but not right exact!)
		\item If $X$ is affine, then the derived functors are 
		\[M\mapsto H^n(X/G,\calL(M))\]
	\end{enumerate}
\end{prop}
\begin{prf}
	The second part looks almost obvious from the Grothendieck spectral sequence. We need, however, 
	that $\calL$ takes injective $G$-modules to acyclic sheaves. Why is that true? Take $M\otimes k[G]$ an 
	injective $G$-module, then we saw that 
	\[\calL(M\otimes k[G])\simeq \pi_\ast\calJ_{X/e}(M).\]
	We need to check that 
	\[H^n(X/G,\pi_\ast\calL_{X/e}(M))\]
	vanishes for all $n\ge 1$. THere is a problem in Hartshorne that gives this to us for an affine morphism between separated (something else) schemes.
\end{prf}
\begin{prop}
	Let $H\le G$. Then 
	\begin{enumerate}
		\item $\calR^n\Ind_H^G(M)\simeq H^n(G/H,\calL_{G/H}(M))$.
		\item If $G/H$ is Noetherian, then $\scrR^n\Ind_H^G(M)=0$ for $n>\dim G/H$.
		\item If $G/H$ is projective, then $\scrR^n\Ind_H^G(M)$ is finitely generated whenever $M$ is finite dimensional.
		\item If $G/H$ is affine, $\Ind_H^G$ is exact.
	\end{enumerate}
\end{prop}
\begin{prf}
	For the first statement, We are just taking higher derived functors of 
	\[H^0(GH,\calL(M))=\calL(M)(G/H)=(M\otimes k[G])^H=\Ind_H^G(M)\]
	where the isomorphisms here are natural. So the derived functors are the same!
\end{prf}
\begin{cor}
	If $H$ is a finite group scheme, then $\Ind_H^G$ is exact.
\end{cor}
\begin{prf}
	When know that $G/H$ is affine and in fact is $\Spec k[G]^H$.
\end{prf}

>>>>

% Document-Specific Macros
\DeclareMathOperator{\SL}{SL}
\DeclareMathOperator{\Dist}{Dist}

@@

\begin{prop}
	\begin{enumerate}
		\item The functor sending 
		\[M\mapsto H^0(X/G,\calL(M))=\Gamma(\calL(M))\]
		is left exact (but not right exact!)
		\item If $X$ is affine, then the derived functors are 
		\[M\mapsto H^n(X/G,\calL(M))\]
	\end{enumerate}
\end{prop}
\begin{prf}
	The second part looks almost obvious from the Grothendieck spectral sequence. We need, however, 
	that $\calL$ takes injective $G$-modules to acyclic sheaves. Why is that true? Take $M\otimes k[G]$ an 
	injective $G$-module, then we saw that 
	\[\calL(M\otimes k[G])\simeq \pi_\ast\calJ_{X/e}(M).\]
	We need to check that 
	\[H^n(X/G,\pi_\ast\calL_{X/e}(M))\]
	vanishes for all $n\ge 1$. THere is a problem in Hartshorne that gives this to us for an affine morphism between separated (something else) schemes.
\end{prf}
\begin{prop}
	Let $H\le G$. Then 
	\begin{enumerate}
		\item $\calR^n\Ind_H^G(M)\simeq H^n(G/H,\calL_{G/H}(M))$.
		\item If $G/H$ is Noetherian, then $\scrR^n\Ind_H^G(M)=0$ for $n>\dim G/H$.
		\item If $G/H$ is projective, then $\scrR^n\Ind_H^G(M)$ is finitely generated whenever $M$ is finite dimensional.
		\item If $G/H$ is affine, $\Ind_H^G$ is exact.
	\end{enumerate}
\end{prop}
\begin{prf}
	For the first statement, We are just taking higher derived functors of 
	\[H^0(GH,\calL(M))=\calL(M)(G/H)=(M\otimes k[G])^H=\Ind_H^G(M)\]
	where the isomorphisms here are natural. So the derived functors are the same!
\end{prf}
\begin{cor}
	If $H$ is a finite group scheme, then $\Ind_H^G$ is exact.
\end{cor}
\begin{prf}
	We know that $G/H$ is affine and in fact is $\Spec k[G]^H$.
\end{prf}
\section{March 2nd, 2020}
Some reminders: we reinterpreted our induction functor in terms of sheaf cohomology. In our particular case, 
we want to restrict to 
\[\calL:\Rep H\to \O_{G/H}\text{-}\mathbf{mod}\]
where $H\le G$ and we are sending $M\mapsto \calL(M)$ where now 
\[\calR^n\Ind_H^G M\simeq H^n(G/H,\calL(M)).\]

\subsection{A Preview}
When we have a Borel subgroup $B\subseteq G$ (e.g. the upper triangulars in $\GL_n$), 
then we get the functor 
\[\Rep B\to \O_{G/B}\text{-}\mathbf{mod}\]
where we send particular representations $\lambda$ to $H^0(G/B,\calL(k_\lambda))$.

\subsection{Leftovers}
\begin{prop}
	$\calL$ is exact.
\end{prop}
\begin{prf}
	It suffices to show that the association of $M$ to $\calL(M)(U)$, for $U$ an affine open, is exact. To 
	prove this, consider $\tilde U=\pi^{-1}\subset X\to U\subset X/G$, which is faithfully flat---that is, the ring map 
	$\pi_\ast:k[U]\to k[\tilde U]$ is. So it suffices to show that the map 
	\[M\mapsto k[\tilde U]\otimes_{k[U]}\calL(M)(U)=k[\tilde U]\otimes_{k[U]}(k[\tilde U]\otimes M)^G\]
	is exact. Recall that we have that $X\times G\simeq X\times_{X/G} X$ via the map $(g,x)\mapsto (g,gx).$
	So via the same map, we get $\tilde U\times_U\tilde U\simeq \tilde U\times G$ where $G$ acts on the right factor.

	This gives us an iso 
	\[k[\tilde U]\otimes_{k[U]}k[\tilde U]\simeq k[\tilde U]\otimes k[G]\]
	and so 
	\[k[\tilde U]\otimes_{k[U]}(k[\tilde U]\otimes M)^G=(k[\tilde U]\otimes_{k[U]}k[\tilde U]\otimes M)^G\simeq (k[\tilde U]\otimes k[G]\otimes M)^G\]
	where in both of the cases below $G$ acts diagonally on the last two tensor factors.

	Then applying the tensor identity, we get that the above is
	\[(k[U]\otimes k[G]\otimes M_{tr})^G\cong k[\tilde U]\otimes M\otimes (k[G])^G\simeq k[\tilde U]\otimes M\]
	which is exact. 
\end{prf}
The next proposition helps us to show that $\calL$ takes injectives to acyclics so we can apply the Grothendieck spectral sequence.
\begin{prop}
	$\calL_{X/G}(M\otimes k[G])\simeq \pi_\ast \calL_{X/e}(M)$.
\end{prop}
\begin{prf}
	It is enough to show that for all $U\subseteq X/G$ open and affine, the two sheaves have the same values. Recall that 
	\[\calL_{X/G}(M\otimes k[G])(U)\cong\Mor(\pi^{-1}(U),M\otimes k[G])^G\cong\Mor(\pi^{-1}U\times G,M)^G\]
	where in the last iso is induced from a kind of $\times$-hom adjunction.

	Then one can define a map on the level of varieties $(x.g)\mapsto (xg,g)$ from $\pi^{-1}U\times G$ to itself where 
	now $\pi^{-1}U$ has the trivial action.  Thus we get 
	\[Mor(\pi^{-1}U_{tr},M)^G\simeq\Mor(\pi^{-1}U,M)=\pi_\ast\calL_{X/e}(M)\]
\end{prf}
\subsection{Special Case}
Say we have $\pi:X\to X/G$ which is locally trivial: for every $U_i\subset X/G$, there is a local section $\sigma_i:U_i\to X$ with 
the properties:
\begin{itemize}
	\item $\cup U_i=X/G$
	\item $\pi\circ\sigma_i=\id_{U_i}$
	\item $\pi^{-1}(U_i)=U_i\times G$
\end{itemize}
A thing we should discuss is the grassmannian. Let's work over a field $k$. Recall that $\operatorname{Gr}(r,n)$ (or similar notation) means the 
$r$ dimensional subspaces in $\A n$. We have a natural action of $\GL_n$ on $\A n$ that induces an action on $\operatorname{Gr}(r,n)$. Then one can show that 
$\GL_n/\operatorname{stab}_{\GL_n}(v)=\Im\pi\simeq\operatorname{Gr}(r,n).$ The stabilizer is a parabolic subgroup. We hope to see more of this 
on Wednesday.

\subsection{More stuff}
\begin{thm}
	If $G$ acts on $X$ and $M\in \Rep G$ is locally trivial (same assumptions everywhere), then if $\pi:X\to X/G$,
	we get that $X\times^G M= X\times M/G\to X/G$ is locally trivial.
\end{thm}
\begin{rmk}
	In particular, if $H\le G$ and if $M\in \Rep H$, the map $G\times^H M\to G/H$ is a vector bundle.
\end{rmk}
\begin{thm}
	Let $H\le G$ and $M\in \Rep H$. Then $G\times ^H M\to G/H$ has the property that $\calL(M)$ is the sheaf of local sections of this vector bundle. That 
	is, if $U\subset G/H$,
	\[\Gamma(U,G\times^H M)\simeq \calL(M)(U)\]
\end{thm}

\section{March 4, 2020}
Today we are talking about 
\subsection{Distribution algebras for an algebraic group}
The idea here is that we know something about representations of algebras and we want to relate
the representations of an algebraic group to the reps of algebras.

One idea is to notice that $\Rep G\simeq \lcomod{k[G]}$ and then dualize to get the modules of the dual of $k[G]$.
This can blow up, of course, so we have to refine this idea.

Let $X$ be an affine scheme and let $k[X]$ be its ring of functions. Let $x\in X(k)$ and let 
\[I_x=\{f\in k[X]|f(x)=0\}\]
Then we define 
\begin{defn}
	$\mu:k[X]\to k$ is a distribution of order $\le n$ supported at $x$ if $\mu(I_x^{n+1})=0$.
	Then $\operatorname{Dist}_n(X,x)\cong \left(k[X]/I_x^{n+1}\right)^\times$ is the collection of all such $\mu$.

	The subset of $\mu$ satisfying $\mu(1)=0$ is $\operatorname{Dist}_n^+(X,x)$ and we write $\operatorname{Dist}(X,x)$ and $\operatorname{Dist}^+(X,x)$
	for the unions over all $n$.
\end{defn}
Notice that $\Dist^+_n(X,x)\cong \left(I_x/I_x^{n+1})\right)^\times$, so 
\[\Dist_1^+(X,x)\simeq (I_x/I_x^2)^\times\simeq T_xX.\]

Now $\Dist(X,x)$ is a $k[X]$-module via ($f,f_1\in k[X],\mu:k[X]\to k$)
\[(f\mu)(f_1)=\mu(ff_1)\]
\begin{rmk}
	If $X$ is a (non-affine) scheme over $k$,
	\[\Dist_n(X,x)\simeq \left(\O_{X,x}/\frakm_{X,x}^{n+1}\right)^\times\]
\end{rmk}
Let's go back to the affine case. This is functorial: if we have a map $\varphi:X\to Y$ and algebra map 
$\varphi^\ast:k[Y]\to k[X]$ that satisfies $(\varphi^\ast)^{-1}(I_x)=I_{\varphi(x)}$. Thus we get a map 
\[d\varphi_x:\Dist_n(X,x)\to \Dist_n(Y,\varphi(x))\]

Also $\Dist(X,x)$ is a cocommutative coalgebra where the diagonal map $\Delta:X\to X\times X$ gives us a map 
\[\Dist(X,x)\to \Dist(X,x)\otimes \Dist(X,x).\]

\begin{ex}
	Let $X=\A1$ and let $x=0$. Recall $k[\A1]\simeq k[T]$. Then $I=I_0=(T)$.
	Therefore 
	\[\Dist_n(\A1,0)\simeq (k[T]/T^{n+1})^\times\]
	and 
	\[\Dist(\A1,0)\simeq \oplus_{r\ge 0}k\gamma_r\]
	where we let $\gamma_r:k[T]\to k$ via $\gamma_r(T^n)=\delta_{n,r}$.

	Therefore $\gamma_1=\frac{\partial}{\partial T}$ and if the characteristic is zero, $\gamma_r=\frac{1}{r!}\frac{\partial}{\partial T^r}$.
	Remember that we have a natural algebra structure from $\A1$ as a group scheme. The coalgebra structure comes from $\gamma_1$ 
	being primitive and extending up to the rest.
\end{ex}

Now let $X=G$ be a (connected!) group scheme and $x=1$. Then $\Dist(G)\eqdef\Dist(G,1)$. 
Then we can get an algebra structure in the following way: if $\mu,\nu:k[G]\to k$,
\[\mu\nu=\mu\bar\otimes\nu\circ\Delta\]
as we usually do with these things.

Now $\Dist G$ is a filtered algebra so that 
\[\Dist_n(G)\Dist_m(G)\subseteq\Dist_{n+m}(G).\]
\begin{rmk}
	In $k[G]$ we have a comultiplication with a special form. This can be read in Jantzen section I.2.4. If $I$ is the 
	augmentation ideal of $k[G]$, and $f\in I$,
	\[\Delta(f)=f\otimes 1+1\otimes f+I\otimes I\]

	Then multiplicativity of $\Delta$ gives us 
	\[\Delta(f_1\cdots f_n)=\prod_1^n(f_i\otimes 1+1\otimes f_n)+\sum_1^{n-1}I^r\otimes I^{n-r}\]
\end{rmk}

If we have $\mu,\nu\in\Dist_n$, then $[\mu,\nu]=\mu\nu-\nu\mu\in\Dist_{n+m-1}(G)$.
Thus we have a filtration $\Dist_n\subseteq \Dist_{n+1}$ and the associated graded ring 
$\oplus \Dist_i/\Dist_{i-1}$ is commutative. This is reminiscent of the universal enveloping algebra of a Lie algebra!
\begin{ex}
	Let $G=\Ga$ again. THen $k[T]=k[\Ga]$ and think of $\gamma_r=\frac{1}{r!}\frac{\partial^r}{\partial T^r}$.
	What is the multiplication? Ends up it is the divided power algebra structure:
	\[\gamma_n\gamma_n=\binom{n+m}{n}\gamma_{n+m}.\]
	How do we see this? Consider 
	\[\Delta(T)=1\otimes T+T\otimes 1\quad\Rightarrow\quad \Delta(T^l)=\sum_0^l\binom{l}{i}T^i\otimes T^{l-i}\]
	so then 
	\[\gamma_n\gamma_m(T^l)=\sum\binom{l}{i}\gamma_n(T^i)\gamma_m(T^{l-i})\]
	where the summands are all zero except when $l=n+m$ and $i=n$.

	In characteristic zero, $\gamma_r$ is honestly $\frac{1}{r!}\gamma_1^r$ so it ends up that this is just isomorphic
	to a polynomial algebra $k[\gamma_1]$. This is not true in positive characteristic, because $\gamma_1^p$ dies, so we get a new 
	generator $\gamma_p$ and so on.
\end{ex}
Now let $\frakg=\Lie G$. Then since $\frakg\simeq\Dist^+_1(G)$ which maps into $\Dist(G)$, the universal property 
of the universal enveloping algebra gets us us the map 
\[\calU(\frakg)\to \Dist(G)\]
\begin{prop}
	In characteristic zero, this map is an isomorphism.
\end{prop}
In positive characteristic, this fails. However we still ahve an injective map 
\[u(\frakg)\hookrightarrow\Dist(G)\]
where $u(\frakg)$ is the usual restricted enveloping algebra.

\section{March 6th, 2020}
Let $M\in\Rep G$ and let $\mu\in \Dist G$. THen define an action of $\Dist G$ on $M$ by 
\[\mu m=(\id\bar\otimes \mu)\circ\Delta_M\]
and we get a map $\Rep G\to\Dist G$. WHen $G$ is an algebraic gorup scheme over $k$,
then the map on homs is an isomorphism. Notice: $\Rep G$ lands in locally finite $\Dist G$ modules, 
but not all locally finite $\Dist G$ algebras come from $G$-mod.

We're going fast today because I asked about $\operatorname{PGL}_n$. Whoops. Gonna just listen.

Check out Jantzen for the development of roots and weights and all. But here are some theorems!
\begin{thm}
	Let $\lambda$ be a dominant weight. Then $H^0(\lambda)\eqdef\Ind_B^G k_\lambda$ is finite dimensional and nonzero, but $H^i(\lambda)=0$ for all $i>0$.
\end{thm}
\begin{thm}
	$H^0(\lambda)$ has a simple socle called $L(\lambda)$. It's obviously also finite dimensional.
\end{thm}
\begin{thm}
	$\{L(\lambda)|\lambda\in X^+\}$ is a complete and irredundant set of representations for $G$.
\end{thm}
